---
title: "N741 Spring 2018 - Homework 7"
subtitle:  "Homework 7 - DUE WED April 11, 2018"
author: "Melinda Higgins"
date: "April 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(error = FALSE)
```

## Homework 7

### Background and Information on HELP Dataset

For homework 7, you will be working with the **HELP** (Health Evaluation and Linkage to Primary Care) Dataset. See complete details posted in Homework 6.

### Variables for Homework 7

For Homework 7, you will focus on these variables from the HELP dataset:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(haven)
helpdata <- haven::read_spss("helpmkh.sav")

h1 <- helpdata %>%
  select(age, female, pss_fr, homeless, 
         pcs, mcs, cesd)

# add dichotomous variable
# to indicate depression for
# people with CESD scores >= 16

h1 <- h1 %>%
  mutate(cesd_gte16 = cesd >= 16)

# change cesd_gte16 LOGIC variable type
# to numeric coded 1=TRUE and 0=FALSE

h1$cesd_gte16 <- as.numeric(h1$cesd_gte16)
attributes(h1$cesd_gte16)$label <- "Indicator of Depression"

# create a function to get the label
# label output from the attributes() function
getlabel <- function(x) attributes(x)$label
# getlabel(sub1$age)

library(purrr)
ldf <- purrr::map_df(h1, getlabel) # this is a 1x15 tibble data.frame
# t(ldf) # transpose for easier reading to a 15x1 single column list

# using knitr to get a table of these
# variable names for Rmarkdown
library(knitr)
knitr::kable(t(ldf),
             col.names = c("Variable Label"),
             caption="Use these variables from HELP dataset for Homework 07")

```

## Homework 7 Assignment

**SETUP** Download and run the "loadHELP.R" `R` script (included in this Github repo [https://github.com/melindahiggins2000/N741Spring2018_Homework6](https://github.com/melindahiggins2000/N741Spring2018_Homework6)) to read in the HELP Dataset "helpmkh.sav". This script also pulls out the variables you need and creates the dichotomous variable for depression `cesd_gte16`.

After running this R script, you will have a data frame called `h1` you can use to do the rest of your analyses. You can also copy this code into your first R markdown code chunk to get you started on Homework 7.

For Homework 7, you will be looking at perceived social support from friends (`pss_fr`) in these subjects. 

```{r}
library(rpart)
library(partykit)
# library(RColorBrewer)
library(reshape2)
# library(NHANES)
# library(mosaic)
# library(faraway) # skip this one

fitcesd <- rpart::rpart(cesd ~ mcs, data = h1)
rpart::printcp(fitcesd) # Display the results
rpart::plotcp(fitcesd) # Visualize cross-validation results
summary(fitcesd) # Detailed summary of fit

# plot tree
plot(fitcesd, uniform = TRUE, compress = FALSE)
text(fitcesd, use.n = TRUE, all = TRUE, cex = 0.5)
```



```{r}
# all vars except the dictomous cesd_gte16
h1a <- h1[,1:7]

# Plot panels for each covariate
h1m <- reshape2::melt(h1a, id.vars = "cesd")
ggplot(h1m, aes(x=cesd, y=value)) +
  geom_point(alpha=0.4)+
  scale_color_brewer(palette="Set2")+
  facet_wrap(~variable, scales="free_y", ncol=3)

# fit a regression tree with all vars
fitall <- rpart::rpart(cesd ~ ., data = h1a)
# Now let's look at fitall
rpart::printcp(fitall) # Display the results
rpart::plotcp(fitall) # Visualize cross-validation results
summary(fitall) # Detailed summary of fit


plot(fitall, uniform = TRUE, compress = FALSE, main = "Regression Tree for CESD Scores from HELP(h1) Data")
text(fitall, use.n = TRUE, all = TRUE, cex = 0.5)

# Prune the tree
pfit <- prune(fitall, cp = fitall$cptable[which.min(fitall$cptable[, "xerror"]), "CP"])

# Plot the pruned tree
plot(pfit, uniform = TRUE, compress = FALSE, main = "Pruned Regression Tree for Ozone")
text(pfit, use.n = TRUE, all = TRUE, cex = 0.5)

```

better graphics with `party`

```{r}
library(party)
fitallp <- party::ctree(cesd ~ ., data = h1a)
plot(fitallp, main = "Conditional Inference Tree for CESD")

```


decision trees - outcome is a group or class

```{r}
# begin with a logistic regression - depressed or not
glm1 <- glm(cesd_gte16 ~ age + female + pss_fr + homeless + 
              pcs + mcs, data = h1)
summary(glm1)

fitk <- rpart::rpart(cesd_gte16 ~ age + female + pss_fr + 
                       homeless + pcs + mcs, 
                     method = "class", data = h1)
class(fitk)
# Display the results
rpart::printcp(fitk)
#Visualize the cross-validation results 
rpart::plotcp(fitk)
# Get a detailed summary of the splits
summary(fitk)
# Plot the tree
plot(fitk, uniform = TRUE, 
     main = "Classification Tree for CESD => 16")
text(fitk, use.n = TRUE, all = TRUE, cex = 0.8)
# Prune the tree
prune_fitk <- prune(fitk, cp = fitk$cptable[which.min(fitk$cptable[, "xerror"]), "CP"])
class(prune_fitk)

# look at cesd_gte16 with ctree from party

h1b <- h1 %>%
  select(age, female, pss_fr, homeless, pcs, mcs, cesd_gte16)
fitallpk <- party::ctree(cesd_gte16 ~ ., data = h1b)
class(fitallpk)
plot(fitallpk, main = "Conditional Inference Tree for CESD => 16")

# Recursive partitioning of CESD on age, 
# female, pss_fr, homeless, pcs, mcs
whoIsDepressed <- rpart::rpart(cesd_gte16 ~ age + female + 
                                 pss_fr + homeless + pcs + mcs,
                               data = h1b, 
                               control = rpart.control(cp = 0.001,
                                                       minbucket = 20))

whoIsDepressed

library(partykit)
# Plot the tree
plot(partykit::as.party(whoIsDepressed))

# EXTRA CREDIT
# Graph as partition
# using the break points shown from the
# conditional tree
ggplot(data = h1b, aes(x = mcs, y = pcs)) +
  geom_count(aes(color = cesd_gte16), alpha = 0.5) +
  geom_vline(xintercept = 50.024) +
  geom_vline(xintercept = 41.164) +
  geom_vline(xintercept = 37.054) +
  geom_segment(x = 37.054, xend = 0, y = 58.164, yend = 58.164) +
  annotate("rect", xmin = 0, xmax = 100, ymin = 0, ymax = 100, fill = "blue", alpha = 0.1) +
  ggtitle("CESD => 16 Partitioned By MCS and PCS - Dark Circles Not Depressed")


```

random forests approach

```{r}
# library(RColorBrewer)
# library(plot3D)
# library(parallel)
library(randomForestSRC)
library(ggRandomForests)

h1a <- as.data.frame(h1a)
set.seed(131)
# Random Forest for the ozone dataset
fitallrf <- randomForestSRC::rfsrc(cesd ~ ., 
                                   data = h1a, ntree = 100, 
                                   tree.err=TRUE)
# view the results
fitallrf
gg_e <- ggRandomForests::gg_error(fitallrf)
plot(gg_e)
# Plot the predicted cesd values
plot(ggRandomForests::gg_rfsrc(fitallrf), alpha = 0.5)
# Plot the VIMP rankins of independent variables
plot(ggRandomForests::gg_vimp(fitallrf))

# Select the variables
varsel_cesd <- randomForestSRC::var.select(fitallrf)
glimpse(varsel_cesd)

# Save the gg_minimal_depth object for later use
gg_md <- ggRandomForests::gg_minimal_depth(varsel_cesd)
# Plot the object
plot(gg_md)

# Plot minimal depth v VIMP
gg_mdVIMP <- ggRandomForests::gg_minimal_vimp(gg_md)
plot(gg_mdVIMP)

#Create the variable dependence object from the random forest
gg_v <- ggRandomForests::gg_variable(fitallrf)

# Use the top ranked minimal depth variables only, plotted in minimal depth rank order
xvar <- gg_md$topvars

# Plot the variable list in a single panel plot
plot(gg_v, xvar = xvar, panel = TRUE, alpha = 0.4) +
  labs(y="Predicted CESD reading", x="")
```

---

**Use R markdown to complete your homework and show all of your code and output in your final report - Turn in a PDF of your report to Canvas. Include a link to your Github repo for Homework 7**

---


